### 2022-08-17 Paper/results-table.tex

method: Geometric, best mrr_ad ** avg: 0.7682305107526883 std: 0.0033724197709617586
method: Geometric, best mrr_ar ** avg: 0.783393817204301 std: 0.002870980798921243
method: Geometric, best mrr_ld ** avg: 0.8963622311827957 std: 0.0038165384708184025
method: Geometric, best mrr_lr ** avg: 0.9112627688172041 std: 0.0073097134335746074
method: Geometric, best mrr_lad ** avg: 0.8920712365591397 std: 0.004478468907650514
method: Geometric, best mrr_lar ** avg: 0.9095463709677419 std: 0.008254095620938519
method: Geometric, best mrr_ard ** avg: 0.7937331989247312 std: 0.002892081483060137
method: Geometric, best mrr_lrd ** avg: 0.9229038978494624 std: 0.005072886474275548
method: Geometric, best mrr_lard ** avg: 0.9214092741935485 std: 0.004504231901153245
method: Geometric, best acc_ad ** avg: 0.6194758064516129 std: 0.005500153720300029
method: Geometric, best acc_ar ** avg: 0.6434274193548386 std: 0.005332041933055135
method: Geometric, best acc_ld ** avg: 0.8202822580645162 std: 0.0057175656281111785
method: Geometric, best acc_lr ** avg: 0.8459677419354839 std: 0.010990399982684721
method: Geometric, best acc_lad ** avg: 0.8108064516129032 std: 0.008065120945062159
method: Geometric, best acc_lar ** avg: 0.8400000000000001 std: 0.014046271275059837
method: Geometric, best acc_ard ** avg: 0.6583870967741936 std: 0.006289030057447286
method: Geometric, best acc_lrd ** avg: 0.8641129032258064 std: 0.008277431297605348
method: Geometric, best acc_lard ** avg: 0.8593548387096774 std: 0.007356633552926914
method: SupCon, best mrr_ad ** avg: 0.7818461021505375 std: 0.0058056222339268986
method: SupCon, best mrr_ar ** avg: 0.7968904569892473 std: 0.005438750540616555
method: SupCon, best mrr_ld ** avg: 0.8903924731182796 std: 0.008784518487631803
method: SupCon, best mrr_lr ** avg: 0.9056451612903225 std: 0.007443727195959732
method: SupCon, best mrr_lad ** avg: 0.8875026881720431 std: 0.0065906156485880715
method: SupCon, best mrr_lar ** avg: 0.9049993279569893 std: 0.006934618496482506
method: SupCon, best mrr_ard ** avg: 0.8120443548387095 std: 0.0038831681475468686
method: SupCon, best mrr_lrd ** avg: 0.919635752688172 std: 0.00421253081123992
method: SupCon, best mrr_lard ** avg: 0.9202607526881721 std: 0.007036465744801571
method: SupCon, best acc_ad ** avg: 0.6417338709677419 std: 0.009174605555378065
method: SupCon, best acc_ar ** avg: 0.6652016129032258 std: 0.010682516325605402
method: SupCon, best acc_ld ** avg: 0.8105241935483869 std: 0.012173413141473797
method: SupCon, best acc_lr ** avg: 0.8364516129032259 std: 0.014036892078731987
method: SupCon, best acc_lad ** avg: 0.8058467741935484 std: 0.011178885534336718
method: SupCon, best acc_lar ** avg: 0.8354435483870969 std: 0.012324733809520844
method: SupCon, best acc_ard ** avg: 0.6869758064516129 std: 0.006569735184573384
method: SupCon, best acc_lrd ** avg: 0.8605645161290323 std: 0.012074307093137743
method: SupCon, best acc_lard ** avg: 0.8581854838709677 std: 0.01293003766775218
method: EMMA, best mrr_ad ** avg: 0.7763420698924731 std: 0.0029124808921936005
method: EMMA, best mrr_ar ** avg: 0.786581317204301 std: 0.006418196740535436
method: EMMA, best mrr_ld ** avg: 0.8986505376344086 std: 0.005024686589459191
method: EMMA, best mrr_lr ** avg: 0.9125739247311827 std: 0.008583791522680621
method: EMMA, best mrr_lad ** avg: 0.8966283602150538 std: 0.0036157579057873954
method: EMMA, best mrr_lar ** avg: 0.9097157258064514 std: 0.00655691760402764
method: EMMA, best mrr_ard ** avg: 0.8032096774193549 std: 0.004465339908648459
method: EMMA, best mrr_lrd ** avg: 0.9270524193548388 std: 0.004950684287669308
method: EMMA, best mrr_lard ** avg: 0.9272177419354838 std: 0.0046582728067974
method: EMMA, best acc_ad ** avg: 0.6353629032258065 std: 0.00531983066789421
method: EMMA, best acc_ar ** avg: 0.6506854838709677 std: 0.010086772331454559
method: EMMA, best acc_ld ** avg: 0.8277822580645161 std: 0.009658415480510912
method: EMMA, best acc_lr ** avg: 0.850725806451613 std: 0.014181056636432354
method: EMMA, best acc_lad ** avg: 0.8216129032258064 std: 0.006363306949864157
method: EMMA, best acc_lar ** avg: 0.8437096774193549 std: 0.012343847760054961
method: EMMA, best acc_ard ** avg: 0.6768951612903227 std: 0.008115363892853094
method: EMMA, best acc_lrd ** avg: 0.8737903225806452 std: 0.007140626974573769
method: EMMA, best acc_lard ** avg: 0.8715322580645161 std: 0.007248425499130212
method: Contrastive, best mrr_ad ** avg: 0.7174408602150538 std: 0.007255402799973243
method: Contrastive, best mrr_ar ** avg: 0.7337029569892473 std: 0.0039224209303724234
method: Contrastive, best mrr_ld ** avg: 0.8972479838709677 std: 0.005357753897962892
method: Contrastive, best mrr_lr ** avg: 0.9082345430107527 std: 0.003749222854686774
method: Contrastive, best mrr_lad ** avg: 0.8912668010752688 std: 0.006122357196224834
method: Contrastive, best mrr_lar ** avg: 0.9026169354838709 std: 0.005789581070670862
method: Contrastive, best mrr_ard ** avg: 0.7496478494623655 std: 0.004373982216482813
method: Contrastive, best mrr_lrd ** avg: 0.9192137096774193 std: 0.00411868564896933
method: Contrastive, best mrr_lard ** avg: 0.9171935483870965 std: 0.005344248282948013
method: Contrastive, best acc_ad ** avg: 0.5482258064516129 std: 0.014024724525739987
method: Contrastive, best acc_ar ** avg: 0.5727016129032259 std: 0.006437231955968572
method: Contrastive, best acc_ld ** avg: 0.82875 std: 0.008802521869277298
method: Contrastive, best acc_lr ** avg: 0.8434677419354838 std: 0.01012939823665928
method: Contrastive, best acc_lad ** avg: 0.8155241935483872 std: 0.00933622313545278
method: Contrastive, best acc_lar ** avg: 0.8326209677419355 std: 0.010162090342290321
method: Contrastive, best acc_ard ** avg: 0.5938306451612902 std: 0.005993840056794196
method: Contrastive, best acc_lrd ** avg: 0.8631048387096774 std: 0.006685527646846855
method: Contrastive, best acc_lard ** avg: 0.8574999999999999 std: 0.008746747574507578