### 2022-10-06 Code/result-analysis/tableACC.tex
Results for TMLR

\begin{tabular}{|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|} 
 \hline 
Methods & speech/depth & speech/RGB & text/depth & text/RGB & text/speech/ \newline depth & text/speech/ \newline RGB & speech/RGB/ \newline depth & text/RGB/ \newline depth & all \\ 
 \hline 
Geometric & 61.95$\pm$0.55 & 64.34$\pm$0.53 & 82.03$\pm$0.57 & 84.6$\pm$1.1 & 81.08$\pm$0.81 & 84.0$\pm$1.4 & 65.84$\pm$0.63 & 86.41$\pm$0.83 & 85.94$\pm$0.74 \\ 
SupCon & \textbf{64.17}$\pm$0.92 & \textbf{66.52}$\pm$1.07 & 81.05$\pm$1.22 & 83.65$\pm$1.4 & 80.58$\pm$1.12 & 83.54$\pm$1.23 & \textbf{68.7}$\pm$0.66 & 86.06$\pm$1.21 & 85.82$\pm$1.29 \\ 
EMMA & 63.54$\pm$0.53 & 65.07$\pm$1.01 & 82.78$\pm$0.97 & \textbf{85.07}$\pm$1.42 & \textbf{82.16}$\pm$0.64 & \textbf{84.37}$\pm$1.23 & 67.69$\pm$0.81 & \textbf{87.38}$\pm$0.71 & \textbf{87.15}$\pm$0.72 \\ 
Contrastive & 54.82$\pm$1.4 & 57.27$\pm$0.64 & \textbf{82.88}$\pm$0.88 & 84.35$\pm$1.01 & 81.55$\pm$0.93 & 83.26$\pm$1.02 & 59.38$\pm$0.6 & 86.31$\pm$0.67 & 85.75$\pm$0.87 \\ 
\hline \end{tabular}