### 2022-10-06 Paper/Figures/unused/results-table-old.tex
Paper submitted to TMLR

\begin{table*}[tb]
    \centering
    \begin{subtable}[t]{1.00\textwidth}
    \centering
    \resizebox{.99\textwidth}{!}{
    \csvautotabular{resultsMRR.csv}
    }
    \end{subtable}
    \vskip\baselineskip
    % \hfill
    \begin{subtable}[t]{1.00\textwidth}
    \centering
    \resizebox{.99\textwidth}{!}{
    \csvautotabular{resultsACC.csv}
    }
    \end{subtable}
\end{table*}
% \todocm{For Table 1 - change all mrr.xx to MRR \textit{xx} for consistency with eqn 8, and all acc.xx to Acc \textit{xx}---e.g., column headers should be 'MRR \textit{td}'. Also make sure the caption edits are correct.}
% \todo[inline]{Make it two subtables, and remove MRR and Acc from the columns, and spell out sr, tsrd, and ...}

% Automatic DataTool Method
\DTLloaddb{resultsMRR}{resultsMRR.csv}
\DTLloaddb{resultsACC}{resultsACC.csv}
\begin{table}[]
    \centering
    \resizebox{1.0\textwidth}{!}{
    \DTLdisplaydb{resultsMRR}
    }
    \resizebox{1.0\textwidth}{!}{
    \DTLdisplaydb{resultsACC}
    }
    \caption{Caption}
    \label{table:quantitative}
\end{table}

% Manually constructed table
\begin{table*}[]
\centering
\resizebox{.97\textwidth}{!}{
\begin{tabular}{lc|ll|lll}
\multicolumn{1}{c}{} &  & \multicolumn{2}{c|}{\textbf{Accuracy}} & \multicolumn{3}{c}{\textbf{Mean Reciprocal Rank}} \\ \hline
\multicolumn{1}{c|}{\textbf{Method}} & \textbf{Modalities} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Speech/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Speech/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/Speech/\\ RGB/Depth\end{tabular}}} \\ \hline
\multicolumn{1}{l|}{EMMA} & TSRD & \multicolumn{1}{l|}{\textbf{0.65}$\pm$0.01} & 0.83$\pm$0.006 & \multicolumn{1}{l|}{\textbf{0.79}$\pm$0.002} & \multicolumn{1}{l|}{0.90$\pm$0.003} & 0.90$\pm$0.004 \\
\multicolumn{1}{l|}{EMMA} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.008 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.90$\pm$0.004} & -- \\
\multicolumn{1}{l|}{\supcon{}} & TSRD & \multicolumn{1}{l|}{0.59$\pm$0.007} & 0.84$\pm$0.008 & \multicolumn{1}{l|}{0.75$\pm$0.004} & \multicolumn{1}{l|}{0.91$\pm$0.006} & 0.90$\pm$0.006 \\
\multicolumn{1}{l|}{\supcon{}} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.006 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.91$\pm$0.004} & -- \\
\multicolumn{1}{l|}{Contrastive Cosine} & TSRD & \multicolumn{1}{l|}{0.40$\pm$0.153} & 0.83$\pm$0.006 & \multicolumn{1}{l|}{0.62$\pm$0.034} & \multicolumn{1}{l|}{0.90$\pm$0.004} & 0.89$\pm$0.005 \\
\multicolumn{1}{l|}{Contrastive Cosine} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.005 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.91$\pm$0.003} & -- \\
\multicolumn{1}{l|}{\ours{}} & TSRD & \multicolumn{1}{l|}{--} & 0.83$\pm$0.008 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.90$\pm$0.005} & -- \\ \hline
\end{tabular}
}
\caption{Average and standard deviation of accuracy and MRR over 3 runs with 3 different random seeds on a held-out test set. These results represent performance when no modalities are ablated, in which case supervised contrastive loss, contrastive loss, and EMMA perform similarly (``--'' is an ablated modality and can't be evaluated). All metrics are from 0 to 1, and higher is better. For 5 objects, a random guess would have MRR of 0.33, and the worst case performance would be always ranking the correct item last which gives us 0.2. All models are trained using text as anchor. The batch size is 64 and optimizer is SGD for all experiments except the triplet loss, where batch size is 1 and optimizer is Adam.
%In the Triplet experiment, RGB and Depth embeddings are concatenated to form a single vision embedding since triplet loss cannot be used for more than two modalities.
} 
\label{table:quantitative}
\end{table*}
% \todo[inline]{This table is not referred to until sec:results, which is waaay down from where it's introduced, and it takes some parsing to figure out what's going on. Either add a paragraph under it that starts with something like "Table 1 shows high-level results, in which..." or just move the table to the results section.}

% %Edward moved table above where refernced b/c LaTeX plays games with double-wide objects
% \begin{table*}[bth]
% \centering
% \begin{tabular}{c|c|c|c|c|c|c}
% \toprule
% \textbf{Method} & \textbf{Modalities} & \textbf{acc.srd} & \textbf{acc.trd} & \textbf{MRR.srd} & \textbf{MRR.trd} & \textbf{MRR.tsrd} \\ %\hline
% \midrule
% EMMA (Text) & 4 & \textbf{0.65}$\pm$0.01 & 0.83$\pm$0.006 & \textbf{0.79}$\pm$0.002 & 0.90$\pm$0.003 & 0.90$\pm$0.004 \\
% EMMA (Text) & 3 & - & 0.84$\pm$0.008 & - & 0.90$\pm$0.004 & - \\
% % Simple MMA (Text)  & 3 & ?$\pm$? & ?$\pm$? & ? & ? \\
% % Simple MMA (Text) & 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
% % Simple MMA (Speech) & 3 & ?$\pm$? & ?$\pm$? & ? & ?  \\
% DONE Supervised Contrastive (Text)& 4 & 0.59$\pm$0.007 & 0.84$\pm$0.008 & 0.75$\pm$0.004 & 0.91$\pm$0.006 & 0.90$\pm$0.006  \\
% TODO Supervised Contrastive (Text)& 3 & - & 0.84$\pm$0.006 & - & 0.91$\pm$0.004 & -\\
% DONE Contrastive Cosine (Text) & 4 & 0.40$\pm$0.153 & 0.83$\pm$0.006 & 0.62$\pm$0.034  & 0.90$\pm$0.004 & 0.89$\pm$0.005\\
% TODO Contrastive Cosine (Text) & 3 & - & 0.84$\pm$0.005 & -  & 0.91$\pm$0.003 & -\\
% PROJ Triplet (Text) & 3 & - & 0.83$\pm$0.008 & - & 0.90$\pm$0.005 & -  \\
% % DONE EMMA (Speech) & 3 & ?$\pm$? & ?$\pm$? & - & - \\
% % DONE EMMA (Speech) & 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
% % DONE Supervised Contrastive (Speech)& 3 & ?$\pm$? & ?$\pm$? & - & -\\
% % DONE Supervised Contrastive (Speech)& 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
% % TODO Contrastive Cosine (Speech) & 3 & ?$\pm$? & ?$\pm$?  & - & -\\
% % TODO Contrastive Cosine (Speech) & 4 & ?$\pm$? & ?$\pm$?  & ? & ?\\
% % DONE Triplet (Speech) & 3 & ?$\pm$? & ?$\pm$? & - & -  \\

% % Full MMA (Text) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
% % Simple MMA (Text) w/ neg Adam & 3  & 0.7949$\pm$? & 0.8778$\pm$? \\
% % Simple MMA (Text) w/ neg SGD & 3 & 0.7905$\pm$? & 0.8821$\pm$? \\
% % Triplet (Text) w/ neg & 3 & 0.7497$\pm$? & 0.8522$\pm$? \\
% % Full MMA (Speech) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
% % Simple MMA (Speech) w/ neg & 3 & 0.7949$\pm$? & 0.8778$\pm$? \\
% % Contrastive Cosine (Text) w/ neg & 3 & 0.7894$\pm$? & 0.8788$\pm$?  & ? & ?\\
% % Contrastive (Speech) w/ neg & 3 & ?$\pm$? & ?$\pm$? \\
% % Triplet (Speech) w/ neg & 3 & 0.7497$\pm$? & 0.8522$\pm$? \\
% \bottomrule
% \end{tabular}
% \caption{\label{table:quantitative}Average and standard deviation of accuracy and MRR over 3 runs with 3 different random seeds on a held-out test set. These results represent performance when no modalities are ablated, in which case supervised contrastive loss, contrastive loss, and EMMA perform similarly. All metrics are from 0 to 1, and higher is better. For 5 objects, a random guess would have MRR of 0.33, and the worst case performance would be always ranking the correct item last which gives us 0.2. All models are trained using text as anchor. The batch size is 64 and optimizer is SGD for all experiments except the triplet loss, where batch size is 1 and optimizer is Adam.
% % \todokdinline{maybe adding results on cropped and raw dataset as well. Also results using object class instead of negative sampling}
% }
% \end{table*}

method: Geometric, best mrr_ad ** avg: 0.7682305107526883 std: 0.0033724197709617586
method: Geometric, best mrr_ar ** avg: 0.783393817204301 std: 0.002870980798921243
method: Geometric, best mrr_ld ** avg: 0.8963622311827957 std: 0.0038165384708184025
method: Geometric, best mrr_lr ** avg: 0.9112627688172041 std: 0.0073097134335746074
method: Geometric, best mrr_lad ** avg: 0.8920712365591397 std: 0.004478468907650514
method: Geometric, best mrr_lar ** avg: 0.9095463709677419 std: 0.008254095620938519
method: Geometric, best mrr_ard ** avg: 0.7937331989247312 std: 0.002892081483060137
method: Geometric, best mrr_lrd ** avg: 0.9229038978494624 std: 0.005072886474275548
method: Geometric, best mrr_lard ** avg: 0.9214092741935485 std: 0.004504231901153245
method: Geometric, best acc_ad ** avg: 0.6194758064516129 std: 0.005500153720300029
method: Geometric, best acc_ar ** avg: 0.6434274193548386 std: 0.005332041933055135
method: Geometric, best acc_ld ** avg: 0.8202822580645162 std: 0.0057175656281111785
method: Geometric, best acc_lr ** avg: 0.8459677419354839 std: 0.010990399982684721
method: Geometric, best acc_lad ** avg: 0.8108064516129032 std: 0.008065120945062159
method: Geometric, best acc_lar ** avg: 0.8400000000000001 std: 0.014046271275059837
method: Geometric, best acc_ard ** avg: 0.6583870967741936 std: 0.006289030057447286
method: Geometric, best acc_lrd ** avg: 0.8641129032258064 std: 0.008277431297605348
method: Geometric, best acc_lard ** avg: 0.8593548387096774 std: 0.007356633552926914
method: SupCon, best mrr_ad ** avg: 0.7818461021505375 std: 0.0058056222339268986
method: SupCon, best mrr_ar ** avg: 0.7968904569892473 std: 0.005438750540616555
method: SupCon, best mrr_ld ** avg: 0.8903924731182796 std: 0.008784518487631803
method: SupCon, best mrr_lr ** avg: 0.9056451612903225 std: 0.007443727195959732
method: SupCon, best mrr_lad ** avg: 0.8875026881720431 std: 0.0065906156485880715
method: SupCon, best mrr_lar ** avg: 0.9049993279569893 std: 0.006934618496482506
method: SupCon, best mrr_ard ** avg: 0.8120443548387095 std: 0.0038831681475468686
method: SupCon, best mrr_lrd ** avg: 0.919635752688172 std: 0.00421253081123992
method: SupCon, best mrr_lard ** avg: 0.9202607526881721 std: 0.007036465744801571
method: SupCon, best acc_ad ** avg: 0.6417338709677419 std: 0.009174605555378065
method: SupCon, best acc_ar ** avg: 0.6652016129032258 std: 0.010682516325605402
method: SupCon, best acc_ld ** avg: 0.8105241935483869 std: 0.012173413141473797
method: SupCon, best acc_lr ** avg: 0.8364516129032259 std: 0.014036892078731987
method: SupCon, best acc_lad ** avg: 0.8058467741935484 std: 0.011178885534336718
method: SupCon, best acc_lar ** avg: 0.8354435483870969 std: 0.012324733809520844
method: SupCon, best acc_ard ** avg: 0.6869758064516129 std: 0.006569735184573384
method: SupCon, best acc_lrd ** avg: 0.8605645161290323 std: 0.012074307093137743
method: SupCon, best acc_lard ** avg: 0.8581854838709677 std: 0.01293003766775218
method: EMMA, best mrr_ad ** avg: 0.7763420698924731 std: 0.0029124808921936005
method: EMMA, best mrr_ar ** avg: 0.786581317204301 std: 0.006418196740535436
method: EMMA, best mrr_ld ** avg: 0.8986505376344086 std: 0.005024686589459191
method: EMMA, best mrr_lr ** avg: 0.9125739247311827 std: 0.008583791522680621
method: EMMA, best mrr_lad ** avg: 0.8966283602150538 std: 0.0036157579057873954
method: EMMA, best mrr_lar ** avg: 0.9097157258064514 std: 0.00655691760402764
method: EMMA, best mrr_ard ** avg: 0.8032096774193549 std: 0.004465339908648459
method: EMMA, best mrr_lrd ** avg: 0.9270524193548388 std: 0.004950684287669308
method: EMMA, best mrr_lard ** avg: 0.9272177419354838 std: 0.0046582728067974
method: EMMA, best acc_ad ** avg: 0.6353629032258065 std: 0.00531983066789421
method: EMMA, best acc_ar ** avg: 0.6506854838709677 std: 0.010086772331454559
method: EMMA, best acc_ld ** avg: 0.8277822580645161 std: 0.009658415480510912
method: EMMA, best acc_lr ** avg: 0.850725806451613 std: 0.014181056636432354
method: EMMA, best acc_lad ** avg: 0.8216129032258064 std: 0.006363306949864157
method: EMMA, best acc_lar ** avg: 0.8437096774193549 std: 0.012343847760054961
method: EMMA, best acc_ard ** avg: 0.6768951612903227 std: 0.008115363892853094
method: EMMA, best acc_lrd ** avg: 0.8737903225806452 std: 0.007140626974573769
method: EMMA, best acc_lard ** avg: 0.8715322580645161 std: 0.007248425499130212
method: Contrastive, best mrr_ad ** avg: 0.7174408602150538 std: 0.007255402799973243
method: Contrastive, best mrr_ar ** avg: 0.7337029569892473 std: 0.0039224209303724234
method: Contrastive, best mrr_ld ** avg: 0.8972479838709677 std: 0.005357753897962892
method: Contrastive, best mrr_lr ** avg: 0.9082345430107527 std: 0.003749222854686774
method: Contrastive, best mrr_lad ** avg: 0.8912668010752688 std: 0.006122357196224834
method: Contrastive, best mrr_lar ** avg: 0.9026169354838709 std: 0.005789581070670862
method: Contrastive, best mrr_ard ** avg: 0.7496478494623655 std: 0.004373982216482813
method: Contrastive, best mrr_lrd ** avg: 0.9192137096774193 std: 0.00411868564896933
method: Contrastive, best mrr_lard ** avg: 0.9171935483870965 std: 0.005344248282948013
method: Contrastive, best acc_ad ** avg: 0.5482258064516129 std: 0.014024724525739987
method: Contrastive, best acc_ar ** avg: 0.5727016129032259 std: 0.006437231955968572
method: Contrastive, best acc_ld ** avg: 0.82875 std: 0.008802521869277298
method: Contrastive, best acc_lr ** avg: 0.8434677419354838 std: 0.01012939823665928
method: Contrastive, best acc_lad ** avg: 0.8155241935483872 std: 0.00933622313545278
method: Contrastive, best acc_lar ** avg: 0.8326209677419355 std: 0.010162090342290321
method: Contrastive, best acc_ard ** avg: 0.5938306451612902 std: 0.005993840056794196
method: Contrastive, best acc_lrd ** avg: 0.8631048387096774 std: 0.006685527646846855
method: Contrastive, best acc_lard ** avg: 0.8574999999999999 std: 0.008746747574507578