### 2022-10-06 Paper/tableMRR.tex

\begin{tabular}{|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|p{{2cm}}|} 
 \hline 
Methods & speech/depth & speech/RGB & text/depth & text/RGB & text/speech/ \newline depth & text/speech/ \newline RGB & speech/RGB/ \newline depth & text/RGB/ \newline depth & all \\ 
 \hline 
Geometric & 76.82$\pm$0.34 & 78.34$\pm$0.29 & 89.64$\pm$0.38 & 91.13$\pm$0.73 & 89.21$\pm$0.45 & 90.95$\pm$0.83 & 79.37$\pm$0.29 & 92.29$\pm$0.51 & 92.14$\pm$0.45 \\ 
SupCon & \textbf{78.18}$\pm$0.58 & \textbf{79.69}$\pm$0.54 & 89.04$\pm$0.88 & 90.56$\pm$0.74 & 88.75$\pm$0.66 & 90.5$\pm$0.69 & \textbf{81.2}$\pm$0.39 & 91.96$\pm$0.42 & 92.03$\pm$0.7 \\ 
EMMA & 77.63$\pm$0.29 & 78.66$\pm$0.64 & \textbf{89.87}$\pm$0.5 & \textbf{91.26}$\pm$0.86 & \textbf{89.66}$\pm$0.36 & \textbf{90.97}$\pm$0.66 & 80.32$\pm$0.45 & \textbf{92.71}$\pm$0.5 & \textbf{92.72}$\pm$0.47 \\ 
Contrastive & 71.74$\pm$0.73 & 73.37$\pm$0.39 & 89.72$\pm$0.54 & 90.82$\pm$0.37 & 89.13$\pm$0.61 & 90.26$\pm$0.58 & 74.96$\pm$0.44 & 91.92$\pm$0.41 & 91.72$\pm$0.53 \\ 
\hline \end{tabular}